---
title: "Is an Automatic transmission better than a manual transmission"
author: "preethi"
date: "11/20/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Executive Summary

Here at *MotorTrend*, we want to answer to very simple questions  

1) Is an automatic or manual transmission better for MPG
2) Can I quantify the MPG difference between automatic and manual transmissions

We will be using the mtcars dataset to answer these questions  

## MTCARS dataset

We will begin with describing the mtcars dataset. This data was extracted from the 1974 *MotorTrend* magazine and comprises of fuel consumption and 10 other aspects of automobile design (mpg, number of cylinders, displacement, transmission, engine etc...) for 32 automobiles

We can look at the first few rows of this dataset now

```{r mtcars}
summary(mtcars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
